---
title: "Supplementary Material: Drift effects on the multivariate floral phenotype of *Calceolaria polyrhiza* during a postglacial expansion in Patagonia"
author: C.C. Maubecin, A. Cosacov, J. Fornoni, A. N. Sérsic & S. Benitez-Vieyra.
output: pdf_document
---

This is a pdf version of an R Markdown, original routines can be downloaded from https://github.com/santiagombv/random.skewers   
Random skewers is a method that compares the evolutionary response to selection of two variance-covariance matrices, as a measure of matrix similarity. A large number of random selection vectors, called random skewers, is simultaneously applied to both matrices and their response to selection vectors are compared. Given any fixed selection vector, if two matrices are similar, the orientation and magnitude of the response vectors should also be similar. We followed the description of the method in Marroig and Cheverud (2007) to build the present routine.  

### The function *random.skewers*  

```{r}
# Arguments.
# G1 and G2 are variance-covariance matrices.
# rep is the number of random skewers.

# Output.
# The output is a two-column data frame. The column ratio is the response vectors'
# length ratio between G1 nd G2.The column correlation contains measures of the
# co-linearity of the response vectors in the multivariate morphometric space, equal
# to the cosine of the angle between response vectors.

random.skewers<-function(G1, G2, rep){
  N1 <- ncol(G1)
  N2 <- ncol(G2)
  if(N1 != N2) stop("G1 and G2 have different dimensions") else N <- N1
  rd.sk <- function(G1, G2){
    RS1 <- runif(n=N)		
    sign <- c(-1,1)
    RS2 <- vector(length=N)
    for (i in 1:N) RS2[i] <- sample(sign, size=1)*RS1[i]
    RS3 <- RS2/sqrt((t(RS2)%*%RS2))
    dz1 <- G1%*%RS3
    dz2 <- G2%*%RS3
    res <- c(sum(dz1^2)/sum(dz2^2), 
             (t(dz1)%*%dz2)/sqrt((t(dz1)%*%dz1)*(t(dz2)%*%dz2)))
    return(res)
  }
  RES <- data.frame(t(replicate(rep, rd.sk(G1=G1, G2=G2))))
  colnames(RES) <- c("ratio", "correlation")
  return(RES)
}

```

### The function *sim.vec.cor*
```{r}
# Arguments.
# k is the number of traits in the random skewer analysis.
# rep is the number of simmulated vector pairs.

# Output.
# A vector of length rep containing the correlations between random vectors. 

sim.vec.cor <- function(k, rep){
  vec.cor <- function(k){
     RS1 <- runif(n=k)*sample(c(-1,1), size = k, replace = TRUE)
     RS1 <- RS1/sqrt((t(RS1)%*%RS1)) 
     RS2 <- runif(n=k)*sample(c(-1,1), size = k, replace = TRUE)
     RS2 <- RS2/sqrt((t(RS2)%*%RS2))
     ang <- acos(t(RS1)%*%RS2/sqrt((t(RS1)%*%RS1)*(t(RS2)%*%RS2)))
     res <- ifelse(ang > pi/2, cos(pi-ang), cos(ang))
     return(res)
  }
  replicate(rep, vec.cor(k))
}

```

### Worked example.
```{r, eval = FALSE}
# Given two covariance matrices, M1 and M2.
M1 <- matrix(c(0.011, 0.003, 0.009, 0.008, 0.002, 0.005,
               0.003, 0.010, 0.004, 0.006, 0.002, 0.006,
               0.009, 0.004, 0.021, 0.007, 0.001, 0.008,
               0.008, 0.006, 0.007, 0.054, -0.018, 0.007,
               0.002, 0.002, 0.001, -0.018, 0.026, -0.001,
               0.005, 0.006, 0.008, 0.007, -0.001, 0.010), 6, 6)

M2 <- matrix(c(0.008, -0.001, 0.008, -0.003, 0.001, 0.001,
               -0.001, 0.014, 0.003, -0.001, 0.002, -0.004,
               0.008, 0.003, 0.021, 0.002, 0.002, 0.001,
               -0.003, -0.001, 0.002, 0.013, 0.003, -0.001,
               0.001, 0.002, 0.002, 0.003, 0.012, 0.001,
               0.001, -0.004, 0.001, -0.001, 0.001, 0.007), 6, 6)

# Create random skewers.
RS <- random.skewers(G1 = M1, G2 = M2, rep = 1000)

# Obtain mean response vectors' length ratio.
mean(RS$ratio)

# Obtain mean response vectors' correlation.
mean(RS$correlation)

# Estimate the 95% percentile of the expected correlation between random vectors.
SC <- sim.vec.cor(k = 6, rep = 1000)
quantile(SC, probs = 0.95)

```

###References
Cheverud, J.M. & Marroig, G. 2007. Comparing covariance matrices: random skewers method compared to  the common principal components model. *Genet. Molec. Biol.* 30: 461-469.

